---
title: 'Reproducible Research: Peer Assessment 1'
output:
  html_document:
    keep_md: yes
  pdf_document: default
---

- Loading and preprocessing the data
```{r load_data}
x <- read.csv(unz("activity.zip", "activity.csv"))
library(ggplot2)
summary(x)
```

##What is mean total number of steps taken per day?  

```{r total_steps_by_date}
#Calc total steps by date
ts_bydate <- aggregate(x$steps, list(x$date), sum, na.rm=TRUE)
names(ts_bydate) <- c("Date", "Steps")
qplot(ts_bydate$Date, ts_bydate$Steps)
```
 
*The mean number of steps taken per day is
`r as.integer(mean(ts_bydate$Steps))` and the median number of steps taken per day is
`r as.integer(median(ts_bydate$Steps))`.*

##What is the average daily activity pattern?

```{r total_steps_by_interval}
#Calc average steps by interval
ts_byint <- aggregate(x$steps, list(x$interval), mean, na.rm=TRUE)
names(ts_byint) <- c("TimeInterval", "AvgSteps")
#Present histogram of average steps by interval
qplot(ts_byint$TimeInterval, ts_byint$AvgSteps)
```

- The chart above displays the average daily activity pattern across 5 minute intervals.

## Imputing missing values

The total number of rows with NAs: `r sum(is.na(x$steps))`

```{r impute missing data}
x$intervalmean <- ts_byint$AvgSteps
x$imputedsteps <- x$steps
x$imputedsteps[is.na(x$imputedsteps)] <- x$intervalmean

imp_bydate <- aggregate(x$imputedsteps, list(x$date), sum, na.rm=TRUE)
names(imp_bydate) <- c("Date", "Steps")
```

*After missing value replacement, the mean number of steps taken per day is
`r as.integer(mean(imp_bydate$Steps))` and the median number of steps taken per day is
`r as.integer(median(imp_bydate$Steps))`.*



## Are there differences in activity patterns between weekdays and weekends?

```{r weekdays_weekends}
x$day <- weekdays(as.Date(x$date))
x$weekday <- ifelse(x$day == "Saturday", "Weekend", ifelse(x$day == "Sunday", "Weekend", "Weekday"))

weekend_data <- subset(x, weekday == "Weekend")
weekday_data <- subset(x, weekday == "Weekday")

ts_weekday <- aggregate(weekday_data$imputedsteps, list(weekday_data$interval), mean, na.rm=TRUE)
ts_weekend <- aggregate(weekend_data$imputedsteps, list(weekend_data$interval), mean, na.rm=TRUE)
names(ts_weekday) <- c("TimeInterval", "AvgSteps")
names(ts_weekend) <- c("TimeInterval", "AvgSteps")

ggplot(weekday_data, aes(interval, imputedsteps)) + geom_line(colour="blue") + xlab("Interval") + ylab("Average Steps")
ggplot(weekend_data, aes(interval, imputedsteps)) + geom_line(colour="red") + xlab("Interval") + ylab("Average Steps")
```